---
title: "HW2"
output: github_document
Date: "2022-09-24"
---

```{r, message=FALSE}
library(tidyverse)
```

# Problem 1 

## Importing data
```{r}
data("penguins", package = "palmerpenguins")
penguins_df = janitor::clean_names(penguins)
```

## Take a look at the data
```{r}
penguins_df
```
```{r, results='hide'}
skimr::skim(penguins_df)
```
The penguins data frame has "species," "island," and "sex" factor values, and "bill_length_mm," "bill_depth_mm," "flipper_length_mm," 
"body_mass_g," and "year" numeric variables. There are `r nrow(penguins_df)` rows, and `r ncol(penguins_df)` columns, and the mean flipper length is `r mean(penguins_df$flipper_length_mm, na.rm = TRUE)` mm. 

## Scatter plot of flipper_length_mm (y) vs bill_length_mm (x)
```{r}
ggplot(penguins_df, aes(x = bill_length_mm, y = flipper_length_mm), group = species) + 
  geom_point(aes(color = species), na.rm = TRUE)
ggsave(file = "problem1_scatterplot.pdf")
```
# Problem 2 

## A random sample size of 10 from a standard Normal distribution

```{r}
samp <- rnorm(10)
log <- samp > 0

problem2_df <- tibble(
  vec_norm = samp,
  vec_log = log,
  vec_char = (c("my","name","is","Soung","and","nice","to","meet","you","all")),
  vec_factor = factor(c("low","low","high","high","medium","low","medium","low","high","medium"))
)
```
## Taking the mean of each variable
```{r}
mean(problem2_df$vec_norm)
mean(problem2_df$vec_log)
mean(problem2_df$vec_char)
mean(problem2_df$vec_factor)
```
We cannot calculate the mean value of the character and factor vectors. 

## Converting the logical, character, and factor variables to numeric variables
```{r, results = 'hide'}
problem2_df$vec_norm <- as.numeric(problem2_df$vec_norm)
problem2_df$vec_log <- as.numeric(problem2_df$vec_log)
problem2_df$vec_char <- as.numeric(problem2_df$vec_char)
problem2_df$vec_factor <- as.numeric(problem2_df$vec_factor)
mean(problem2_df$vec_norm)
mean(problem2_df$vec_log)
mean(problem2_df$vec_char)
mean(problem2_df$vec_factor)
```
The logical and factor variables become numerical variables while the character variables don't. We can see that in the background, R assigns and stores the logical and factor variables as integers. However, we are unable to calculate the mean directly for the factor variables as R treats the integers as levels in a categorical variable. However, after converting the factor variables to numeric variable, we can calculate the mean. 